---
title: "About"
author: "Teddy Landis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

For my project, I am looking at the effect first-time voting type has on future voting behaviors. If someone's first possible vote (i.e. the first possible election they can vote in after they turn 18), is cast through an absentee ballot, does that affect their likelihood of voting in future elections?

My data comes from two datasets provided by the [North Carolina Secretary of State](https://www.ncsbe.gov/Public-Records-Data-Info/Election-Results-Data) . I am using data from North Carolina because they provide both voter registration records and voter history for free on their website. I will design my project using the datasets for Alamance county, but once my project is built I will use the Statewide datasets. The statewide dataset is very large and slows my computer - I would rather wait until the end of my project to use it. Instructions on how to access this data are available in README.

Dataset 1 is alamance_hist (ncvhis1.txt). It is a .txt file that I downloaded from the North Carolina Secretary of State. It contains records for each vote cast. Dataset 2 is almanace_reg (ncvoter1.txt) and it contains each voter's voter registration record. Both contain a "voter registration number" which I can use to merge the datasets. The statewide datasets are identical but include more observations.

My repository can be found [here](https://github.com/tlandis212/final_project).

*Milestone 4 Update*: I further cleaned up my datafile. It is now smaller and only includes essential variables.


```{r merge, echo = FALSE}
# I didn't include the dataset in my repository because the file are too large.
# Information on how to download my datasets can be found in my README.md. In a
# Piazza post, it was suggested this was a viable alternative to using GitHub
# LFS.

# Load in both datasets

alamance_hist <- read.delim("raw-data/ncvhis1.txt")
alamance_reg <- read.delim("raw-data/ncvoter1.txt")

# Use full join to merge both datasets

alamance_full_voter <- full_join(alamance_hist, alamance_reg, by=c("voter_reg_num"))

# Reduce the number of variables by eliminating unwanted variables and filter to
# only include voters who would be voting for the first time in 2008 or
# post-2008 elections

alamance_full_voter_reduced <- alamance_full_voter %>% 
  select(county_id.x:voted_party_cd, status_cd:absent_ind, birth_year, birth_state) %>% 
  filter(birth_year >= 1990)

```