---
title: "About"
author: "Teddy Landis and Liz Hoveland"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gt)
library(lubridate)

# I didn't include the dataset in my repository because the file are too large.
# Information on how to download my datasets can be found in my README.md. In a
# Piazza post, it was suggested this was a viable alternative to using GitHub
# LFS.

# Load in both datasets

alamance_hist <- read.delim("raw-data/ncvhis1.txt")
alamance_reg <- read.delim("raw-data/ncvoter1.txt")

# Use full join to merge both datasets

alamance_full_voter <- full_join(alamance_hist, alamance_reg, by=c("voter_reg_num"))
```

For my project, I am looking at the effect first-time voting type has on future voting behaviors. If someone's first possible vote (i.e. the first possible election they can vote in after they turn 18), is cast through an absentee ballot, does that affect their likelihood of voting in future elections?

My data comes from two datasets provided by the [North Carolina Secretary of State](https://www.ncsbe.gov/Public-Records-Data-Info/Election-Results-Data) . I am using data from North Carolina because they provide both voter registration records and voter history for free on their website. I will design my project using the datasets for Alamance county, but once my project is built I will use the Statewide datasets. The statewide dataset is very large and slows my computer - I would rather wait until the end of my project to use it. Instructions on how to access this data are available in README.

Dataset 1 is alamance_hist (ncvhis1.txt). It is a .txt file that I downloaded from the North Carolina Secretary of State. It contains records for each vote cast. Dataset 2 is almanace_reg (ncvoter1.txt) and it contains each voter's voter registration record. Both contain a "voter registration number" which I can use to merge the datasets. The statewide datasets are identical but include more observations.

My repository can be found [here](https://github.com/tlandis212/final_project).

*Milestone 4 Update*: I further cleaned up my datafile. It is now smaller and only includes essential variables.

My plan is to create topline descriptions of how many people voted in each election after the previous one.

## Exploring the Dataset

Looking at how the data is formatted, we learn that we're interested in folks who vote "Absentee by Mail". Everyone else can be considered an in person vote. 


```{r summary, echo = FALSE}

# I checked to see what types of voting methods there were. Everything that
# isn't "Absentee by Mail" requires an in-person vote.

alamance_full_voter%>% 
  group_by(voting_method) %>% 
  summarize(count = n()) %>% 
  gt() %>% 
  tab_header(title = "Exploring type of voting in the NC Voter File") %>% 
  cols_label(voting_method = "Voting Method", count = "Count")

```
## Milestone 5

I wanted to see which elections I should focus on. My instinct was to focus on the general elections in 2012, 2014, 2016, and 2018, unless there were any other particularly high turnout elections. My suspicion was correct.

```{r turnout, echo = FALSE}
# Find a list of all elections and their turnout. Due to the high number of
# extremely low turn out local elections, I am limiting this to the top 15. I
# filtered out missing values for election_desc as this filters out voters who
# have never cast a ballot.

turnout <- alamance_full_voter %>% 
  mutate(election_lbl = mdy(election_lbl)) %>% 
  filter(!is.na(election_desc)) %>% 
  count(election_lbl, election_desc) %>% 
  arrange(desc(n)) %>% 
  slice(1:15)

# I plotted turnout for each election, sorted in descending order of turnout. I
# used coord_flip() so the full election names can be shown. Because I'm using
# the count variable from my tibble (rather than doing it through geom_bar()), I
# needed to set stat = "identity".

ggplot(turnout, aes(reorder(election_desc, -n), n)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_classic() +
  labs(title = "Turnout in Alamance, NC by Election", subtitle = "Number of Raw Votes Cast", caption = "Data Source: NC Secretary of State", x = "Election", y = "Votes Cast")
```

  
```{r small_dataset, echo = FALSE }
# Reduce the number of variables by eliminating unwanted variables and filter to
# create our universe of voters: people who turned 18 in 2012 *and* voted that year.

alamance_small <- alamance_full_voter %>% 
  
  # Voters born in 1994 will have turned 18 in 2012
  
  filter(birth_year >= 1994) %>% 
  
  # Create a variable which contains the election, but as a string rather than a
  # date. This is so that we can cast these into variables later when we make
  # our data wider. I'm only relabeling for four elections, since I'm going to
  # filter our dataset to only include these four elections
  
  mutate(election = recode_factor(election_lbl, `11/06/2012` = "gelection_2012",
                `11/04/2014` = "gelection_2014",
                `11/08/2016` = "gelection_2016",
                `11/06/2018` = "gelection_2018")) %>% 
  
  filter(election == "gelection_2012" | election == "gelection_2014" | election == "gelection_2016" | election == "gelection_2018") %>% 

  # Create a dummy variable that signifies if a vote was cast by a mail-in
  # absentee ballot. NC offers several type of absentee voting, but we're only
  # interested in mail-in absentees. All other forms of absentee voting requires
  # in-person precense and we will count this as an in-person vote

  mutate(absentee = ifelse(voting_method == "ABSENTEE BY MAIL", "Absentee", "Not Absentee")) %>% 
  
  # Select the only variables we need now
  
  select(county_id.x:county_desc.x, voter_reg_num, election, absentee) %>% 
  
  # Create a variable for each of the elections and make the value equal to
  # "Absentee", "Not Absentee", "No Vote"
  
  pivot_wider(names_from = election, values_from = absentee, values_fill = list(absentee = "No Vote")) %>% 
  
  # Only include voters who voted in the 2012 election
  filter(gelection_2012 != "No Vote") %>% 
  arrange(voter_reg_num)
```

## Defining the Experiment



```{r ms6, echo = FALSE}

# Use alamance_small for Milestone 6. 

# Identify voters as part of the control or treatment group

treatment <- alamance_small %>% 
 
  # Identify voters as part of the control or treatment group
  
  mutate(treatment = ifelse(gelection_2012 == "Absentee", TRUE, FALSE),
  
         # Create election outcome variables
         
         outcome_2012 = ifelse(gelection_2012 != "No Vote", TRUE, FALSE),
         outcome_2014 = ifelse(gelection_2014 != "No Vote", TRUE, FALSE),
         outcome_2016 = ifelse(gelection_2016 != "No Vote", TRUE, FALSE),
         outcome_2018 = ifelse(gelection_2018 != "No Vote", TRUE, FALSE))


# Print summary statistics of our treatment and control group

treatment %>% 
  group_by(treatment) %>% 
  count() %>% 
  ungroup() %>% 
  gt()

```


```{r}

# Create a new data frame with our summary statistics for the experiment

treatment_outcomes <- treatment %>% 
  
  # Narrow down to the columnms we need
  
  select(treatment, outcome_2012, outcome_2014, outcome_2016, outcome_2018) %>% 
  
  # Pivlot longer so that we have a year column
  
  pivot_longer(-treatment, names_to = "year", names_prefix = "outcome_", values_to = "vote") %>% 
  
  # Prepare to create a table that counts the number of votes for each election,
  # shown by treatment group
  
  group_by(treatment, year, vote) %>% 
  arrange(year) %>% 
  count() %>% 
  
  # Create a column which shows the percentage of each group that voted in each election
  
  mutate(perc_treatment = round(100*ifelse(treatment == TRUE, n/56, n/738), digits = 2))


treatment_outcomes %>% 
  
  # We only need to plot the percentage of those that vote, so we only need data
  # for vote = TRUE
  
  filter(vote == TRUE) %>% 
  ggplot(aes(x= year, y= perc_treatment, fill = treatment)) +
  
  # We already calculated the percentage up above, so we can set stat =
  # "identity"
  
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(title = "General Election Turnout Among Voters in Alamance, NC\nWho Turned 18 and Voted for the First Time in 2012", subtitle = "Voters whose first ballots were cast by Absentee Ballots are in the Treatment group", x = "General Election Year", y = "Percent that Cast Ballots", caption = "Explanation for why 2012 is 100%") +
  
  # Add data labels
  
  geom_text(aes(label = perc_treatment), position = position_dodge(.9), vjust = 2) +
  theme_classic()


# Types of regressions to do Voting conditional on voting in the first. Use glm.
# Look at descriptively who does mail-in voting - if we don't allow this how would this hurt. Density map of where mail-in voting happens. When running state-wide data - let it sit.

# Density map - interactive, hover over?
# Turnout for this cohort over time

# About Panel, model panel, analysis panel - one of which is interactive Regress
# outcome of voting in 2014, 2016, and 2018... Explanatory variable - voting by
# absentee ballot in 2012. Add controls like age race gender. Write out
# equations in statistical models page. Explain what different variables are.
# Appendix with different graphs.
  
```
```

